### 2018-11-15 index.Rmd
changed the repo structure to a bookdown article

---
output:
  pdf_document:
    keep_tex: true
    template: sagej_template.tex
bibliography: ["references.bib"]
biblio-style: apa
link-citations: yes
citation_package: biblatex
---

\runninghead{Smith and Wittkopf}

\title{workingtitle}

\author{Clemens Schmid\affilnum{1}}

\affiliation{\affilnum{1}Institute of Pre- and Protohistoric Archaeology, Kiel University, Kiel, DE}

\corrauth{
Clemens Schmid,
Institut für Ur- und Frühgeschichte,
Christian-Albrechts-Universität,
D - 24098 Kiel.
}

\email{clemens@nevrome.de}

\begin{abstract}
This paper is nice.
\end{abstract}

\keywords{Bronze Age, Burial Traditions, Cultural Evolution, Cultural Transmission}

\maketitle