### 2018-11-25 01_Introduction.Rmd
more love for introduction

# Introduction {#introduction}

This articles employs some of the methods that arose within archaeology's contemplation of *Cultural transmission theory* to trace the development of four major bronze age burial traditions in Europe. The cultural-historical evolution of archaeological cultures and ethnic groups during the european bronze age is complex and it's reconstruction an ongoing endeavour of archaeology. Within this history of constant emergence, displacement and demise many different forms of burial traditions were practiced. Most of them can be distinguished by the criteria *burial construction* and *treatment of the body*. Bronze age burials are either flat graves or part of a burial mound, and they contain either untreated or cremated human remains. Openly available bulk radiocarbon data now allows to reconstruct the trends within the distribution of these complex behaviours to test classical archaeological narratives about their respective prevalence. The comparison of macroregional trends also opens the perspective on long term patterns of interaction intensity on a continental scale.

The spread of ideas in human populations over space and time is a major topic of *Cultural evolution theory* [@creanzaCulturalEvolutionaryTheory2017]. This framework is as least as old as theories of biological evolution and underwent a long and versatile research history from *Evolutionism* and *Social Darwinism* in the 19th and early 20th century, against *Cultural relativism* and back to *Neoevolutionism* up until the ninteen sixties and seventies [@petermannGeschichteEthnologie2004]. In the late seventies the the discussion took another turn away from social structure to the more general observation of human behaviour. Three major theoretical movements emerged in this context: *Evolutionary psychology*, *Human behavioural ecology* and *Dual inheritance theory* [@smithThreeStylesEvolutionary2000]. The latter -- heavily popularized by Richard Dawkin's *Memetics* [@dawkinsSelfishGene1976] -- establishes the notion of culture acting as secondary heredity system in humans [@henrichDualinheritanceTheoryEvolution2007]. Just like in biological evolution there might be selection and mutation within *Social learning* and therefore methods of *Population genetics* could be employed within cultural anthropology -- of course with the much faster pace of cultural exchange in mind. A lot of theoretical and quantitative research arose from this assumptions: e.g. @cavalli-sforzaCulturalTransmissionEvolution1981, @boydCultureEvolutionaryProcess1985,  @teltserEvolutionaryArchaeologyMethodological1995 or @obrienEvolutionaryArchaeologyTheory1996.

Archaeology has a special perspective to offer for Cultural evolution research: It's concerned with questions that can embrace continent size areas and millennia of cultural development [@garveyCurrentPotentialRoles2018]. Archaeological data is mostly comprised of the material remains of human behaviour -- so the ruins of past architecture, individual lost or discarded artefacts and all more inconspicuous by-products of live. We rightfully expect humans to learn handcraft as well als building techniques and traditions  -- both functionally meaningful and seemingly pointless -- from others close to them. The similarity of historically and spatially close material remains can therefore be explained as a result of social interaction [@lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].

In 1978 Robert Dunnell formulated in a contribution a previously discussed and in the following intensive to a distinction that has been taken up: In archaeological contexts and over long periods of time, ideas would have to be and behaviour patterns into two categories function (functions) and form (style). Functional Ideas on how the knowledge of the use of a cereal form, the competence of certain cereal Tools or weapons to manufacture or social recognition of child abundance are directly relevant to selection. This is where the biological evolution, the survival rate in crisis situations, or of demographic development, to favor certain ideas about their carriers. This Ideas are opposed by many others: Newtral traits that are not directly relevant for selection. such as ceramic decorations, traditional costume fashion or Jewelry shape. Survival or reproductive ability of a population are usually not affected by the nature of the geometric patterns on their storage vessels dependent. However, these characteristics pass through a partly breathtakingly fast development and allow archaeologists insights into relative chronology and social structure. According to Dunnell, in the context of the of Evolutionary Archaeology despite its genetically subordinate clout deserve special relevance - and a dedicated set of methods: Dunnell postulates, that functional ideas, taking into account the quality of their adaptation could be modelled, while stylistic Ideas essentially random, stochastic processes (Neutral Transmission) would follow.

Dunnell's distinction fits in well with the other Context of the Dual Inheritance Theory (see Chapter 2.1.4) and has been taken up in a variety of ways, and extended. In reality, however, most of the Neutral variants hard selection criteria not completely sublime: Their creation causes in different Scope of costs in terms of working time and material. Since the distinction between stylistic and functional characteristics are not always obvious. it may be necessary to evaluate the functional usefulness of the of different variants with experiments and other methods of engineering science: The more similar the usefulness of all available variants in a context, the larger the probability that they will become selective relative to each other, and neutral.

Fraser Neimans Article Stylistic Variation in Evolutionary Perspective: Inferences from Decorative Diversity and Interassemblage Distance in Illinois Woodland Ceramic Assemblages explores the new technologies proposed by Dunnell. stochastic processes and develops a simple framework for the simulation of propagation of non-functional ideas. In the process, the essential phenomena drift and flow (at Neiman: intergroup transmission) from population genetics in archaeology and to the interpretation of the Cultural distances between Indian ceramic inventories is used. Neiman's contribution was well received by the experts. very intensively received. Carl can be called Lipos works on ceramics in the Mississippi underflow, the Cultural Transmission Simulation with classical archaeological seriation, Shennans Investigation of the technological development of linear-band ceramics Ceramics in Merzbachtal or Timothy Kohlers analysis of the diversity of forms and its Role as proxy to the prehistoric social fabric in the southwest of the USA, again with ceramics . Also for the preparation of the simulation of this work was his contribution is an essential inspiration (see chapter 4.3.1).

Evolutionary Archaeology has - parallel to the further development of the the science of archaeology - in the past 15 years of dynamism. The latest Developments show that they are more successful than almost any other archaeological faculty the scientific and Ideal for gradual, theoretical further development from one contribution to the next. During Neiman the concept of his simulation from population genetics was borrowed from other methodology for the Preparation and investigation of phylogenetic trees from cladistics to archaeological questions applied . Stylistic Variability is a very promising basis for modelling with regard to cultural (in comparison to spatial) distance. The Bayes' theorem has proven to be a potent statistical tool. to the model analysis and evaluation revealed . Archaeological data sets and their special features for Cultural evolution research is becoming more rigorous today. questioned and analyzed more consciously - beyond the anecdotal evidence.