### 2018-12-01 02_Materials_and_Methods.Rmd
and even more

# Materials and Methods {#materials-and-methods}

The analysis code of this paper is written in the statistics programming language R [@rcoreteamLanguageEnvironmentStatistical2018] and fully reproducible: **doi**.

The dataset used for the analysis in this paper is a subset of the openly available database Radon-B [@kneiselRadonB2013]. It contains ^14^C dates from the European Bronze and Early Iron Age (2300 BC to 500 BC). Radiocarbon dating is one of the major absolute dating techniques for this period [@ramseyRadiocarbonDatingRevolutions2008] and Radon-B (as of December 2018) is a collection of more than 11000 published dates from around 2800 sites from all over europe. Beyond the identifying laboratory number, the actual, uncalibrated date and a literature reference, the ^14^C dates in Radon-B are also documented with context information about the name and location of the dated site (coordinates), the sample material and the the type and cultural attribution of the sampled context. The variable *Feature type* systematically distinguishes for example between settlements, rockshelters and also different kinds of graves: *Grave (cremation), Grave (flat) inhumation*, etc.. This information allows to accurately and precisely determine a moment in space and time, when a grave of a certain type was created. In theory the more radiocarbon dates with this information are available, the better becomes our understanding and reconstruction of burial traditions in prehistoric Europe. Unfortunately the Radon-B dataset has some data quality issues: **TODO**. The data representativity and semantic consequences will be discussed below.

The following major steps of Data preparation were applied. Please see the code for the details of implementation. The value in brackets indicates the number of dates still available after the respective filtering operation.

1. Downloading the Radon-B dataset with the R package c14bazAAR [@schmidC14bazAARDownloadPrepare2018], which allows to access several open radiocarbon archives. (11048 dates)
2. Removing all dates without age information or age information outside of the limits of the Intcal13 calibration curve: 71 - 46401 BP [@reimer_intcal13_2013]. (10956 dates)
3. Calibrating the ^14^C dates with the R package Bchron [@haslettSimpleMonotoneProcess2008], which implements the calibration algorithm with numerical integration.
4. Reducing date selection to all dates whose $2\sigma$ calibrated age range overlap by at least one year with the predefined window of interest: 2200 -- 800 calBC. (7543 dates)
5. Reducing the date selection to all dates taken from graves as indicated by the *Feature type* variable. (2361)
6. Transforming the variable *Feature type* via pattern matching to two new variables *burial type* (inhumation/cremation/unknown) and *burial construction* (mound/flat/unknown).
7. Manually defining a rectangular (EPSG:102013) research area with sufficient information in Central-, Northern and Northwestern Europe based on perceived spatial date distribution (Figure **TODO**).
8. Reducing date selection to dates within the research area. (1894 dates)
9. Removing ^14^C date doubles with tools of c14bazAAR. (1848 dates)
10. Transforming data from the meaning *^14^C dates on graves* to *burial type at a certain time*. This is a more complex process, because **TODO** (1704 graves)
11. Defining macroregions. **TODO**
12. Reducing the date selection to the space covered by the macroregions and attributing the dates to the respective units. (1562 graves)

The resulting dataset **TODO**

- Spatial Distance

- Cultural distance

The SED is defined as:

$$d_{ij}^2 = \sum_{k = 1}^{n} (p_{ik} - p_{jk})^2$$

- $d_{ij}^2$: SED between tqo groups $i$ and $j$
- $k$: trait counter 
- $n$: number of traits in the population
- $p_{ik}$: relative frequency of the $k$th trait in group $i$
- $p_{jk}$: relative frequency of the $k$th trait in group $j$

- Simulation