### 2019-01-27 article/index.Rmd
update of affiliation

---
title: A radiocarbon data based evaluation of Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe
runninghead: Schmid.
author:
- name: Clemens Schmid
  num: 1
address:
- num: 1
  org: Römisch-Germanisches Zentralmuseum, Leibniz-Forschungsinstitut für Archäologie, Mainz, Germany
corrauth: Clemens Schmid, Römisch-Germanisches Zentralmuseum, Leibniz-Forschungsinstitut für Archäologie, D - 55116 Mainz.
email: clemens@nevrome.de
abstract: Petersilie.
keywords: Bronze Age, Burial Traditions, Cultural Evolution, Cultural Transmission
classoption:
  - Afour
  - times
bibliography: ../references.bib
bibliographystyle: apalike
output:
  rticles::sage_article:
    keep_tex: yes
    includes:
        in_header: preamble.tex
---

```{r echo=FALSE}
library(magrittr)

# allow easy dynamic value integration
sf_prep <- "../analysis/data/output_data/sf_prep.txt"
sf_desc <- "../analysis/data/output_data/sf_desc.txt"
r <- txtstorage::read
a <- function(x) {r("regions graves amounts", sf_prep) %>% strsplit(", ") %>% unlist %>% magrittr::extract(grepl(x, .))}
```

```{r child = '01_Introduction.Rmd'}
```

```{r child = '02_Materials_and_Methods.Rmd'}
```

```{r child = '03_Results.Rmd'}
```

```{r child = '04_Discussion.Rmd'}
```

```{r child = '05_Conclusion.Rmd'}
```

```{r child = '06_Remarks.Rmd'}
```