### 2019-02-05 README.Rmd
some more info about the directory structure in the readme

---
output:
  md_document:
    variant: gfm
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE} 
knitr::opts_chunk$set( collapse = TRUE, comment = "#>", fig.path = "README-" )
```

[![Build Status](https://travis-ci.com/nevrome/cultrans.bronzeageburials.article2019.svg?token=vxsQ9RjxoGASGtX4Q8jc&branch=master)](https://travis-ci.com/nevrome/cultrans.bronzeageburials.article2019) [![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-brightgreen.svg)](https://github.com/nevrome/nevrome/cultrans.bronzeageburials.article2019/commits/master)
[![minimal R version](https://img.shields.io/badge/R%3E%3D-3.5.0-brightgreen.svg)](https://cran.r-project.org/)
[![ORCiD](https://img.shields.io/badge/ORCiD-0000--0003--3448--5715-green.svg)](http://orcid.org/0000-0003-3448-5715)

## Research compendium for ‘Evaluating Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe using radiocarbon data’

### Compendium DOI:

<http://dx.doi.org/>…

The files at the URL above will generate the results as found in the publication. The files hosted at <https://github.com/nevrome/cultrans.bronzeageburials.article2019> are the development versions and may have changed since the paper was published

### Author of this repository:

Clemens Schmid (<clemens@nevrome.de>)

### Published in:

Not yet published.

### Abstract:

European Bronze Age archaeology traditionally focusses on two major dimensions to categorise burials -- although there is an immense variability of attendant phenomena within this spectrum -- flat graves versus burial mounds and cremation versus inhumation. These traits are an indispensable ingredient for common archaeological narratives of sociocultural interaction and cultural evolution. This paper presents a quantitative reconstruction of the general trends in the distribution of Bronze Age burial traditions based on bulk radiocarbon data and employs the resulting time series for the estimation of macro-regional cultural distance. Despite the relatively small amount of input data the trend reconstruction fits to established archaeological observations for prehistoric Europe. The comparison of cultural and spatial distance leads to the remarkable result of no significant permanent correlation, which indicates that the spread of the relevant aspects of burial traditions can not be explained with simple diffusion models. Instead a more complex process of cultural transmission has to be considered.

### Keywords:

Bronze Age, Burial Traditions, Cultural Evolution, Cultural Transmission

### Overview of contents:

This repository contains text, code and data for the paper. The paper is rendered via continuous integration and can be downloaded from the [document branch](https://github.com/nevrome/cultrans.bronzeageburials.article2019/tree/document), so
[here](https://github.com/nevrome/cultrans.bronzeageburials.article2019/blob/document/article.pdf).

The `analysis` directory contains `code` and `data` to reproduce the preparations, calculations, simulations and figure renderings in this paper. The `article` directory contains the tex template and the text for the paper in *.Rmd* format.

### How to reproduce:

As the data and code in this repository are complete and self-contained, it can be reproduced with any R environment (\> version 3.5.0). The necessary package dependencies are documented in the `DESCRIPTION` file and can be installed manually or automatically with `devtools::install()`. If it’s not possible any more to construct a working environment with these methods due to technological progress, one can use the Docker image.

A Docker image is a lightweight GNU/Linux virtual computer that can be run as a piece of software on Windows, Linux, and OSX. To capture the complete computational environment used for this project we have a Dockerfile that specifies how to make the Docker image that we developed this project in. The Docker image includes all of the software dependencies needed to run the code in this project, including the data and code itself. To launch the Docker image for this project, first, [install Docker](https://docs.docker.com/installation/) on your computer and download the `.tar` file with the paper image [here](...) (not yet available). At the Docker prompt, you can load and run the image with:

    docker load -i cultransbronze19_docker_image.tar
    docker run -e PASSWORD=cultransbronze19 -dp 8787:8787 --name cultransbronze19 cultransbronze19

This will start a server instance of RStudio. Then open your web browser at localhost:8787 or run `docker-machine ip default` in the shell to find the correct IP address, and log in with rstudio/cultransbronze19. Once logged in, use the Files pane (bottom right) to navigate to the script files. More information about using RStudio in Docker is available at the [Rocker](https://github.com/rocker-org) [wiki](https://github.com/rocker-org/rocker/wiki/Using-the-RStudio-image) pages.

We developed and tested the package on this Docker container, so this is the only platform that we’re confident it works on. It was built and stored with:

    docker build -t cultransbronze19 .
    docker save -o cultransbronze19_docker_image.tar cultransbronze19

### Licenses:

Code: MIT <http://opensource.org/licenses/MIT> year: 2019, copyright holder: Clemens Schmid

Data: Please see the license agreements of Radon-B and Natural Earth data

Text: Please see the license agreements of the Adaptive Behavior journal